### 2011-05-11 paper.tex
finished proof of oracle theorem

\section{Polynomial Time Reductions}

One of the technical details that will arise in many of our results is the need for polynomial time reductions. In this section, we will establish that all $\{\Sigma,\Pi,\bp,\parity\}$-constructible complexity classes are closed under polynomial time reductions. In particular, we need.

\begin{prop}
Let $C$ be a $\{\Sigma,\Pi,\bp,\parity\}$-constructible complexity class and let $T$ be a mapping $\{0,1\}^n \to \{0,1\}^m$ defined by a polynomial time Turing machine. Then if $f \in C$, $f\circ T \in C$.
\end{prop}

This follows from the following four lemmas:

\begin{lemma}
If $C$ is a complexity class closed under polynomial time reductions. Then $\Sigma \cdot C$ is closed under polynomial time reductions.
\end{lemma}
\begin{proof}
\end{proof}
We wish to show that if $f \in \Sigma \cdot C$, then $f \circ T \in \Sigma \cdot C$. We have that 
\begin{lemma}
If $C$ is a complexity class closed under polynomial time reductions. Then $\Pi \cdot C$ is closed under polynomial time reductions.
\end{lemma}
\begin{proof}

\end{proof}

\begin{lemma}
If $C$ is a complexity class closed under polynomial time reductions. Then $\bp \cdot C$ is closed under polynomial time reductions.
\end{lemma}
\begin{proof}

\end{proof}

\begin{lemma}
If $C$ is a complexity class closed under polynomial time reductions. Then $\parity \cdot C$ is closed under polynomial time reductions.
\end{lemma}
\begin{proof}

\end{proof}

@@

\begin{theorem}\label{amplify}
Let $C$ be a \emph{$\{\Sigma,\Pi,\bp,\parity\}$}-constructible complexity class. Then \emph{$\bp \cdot C = \strongbp \cdot C$}.
\end{theorem}
Initially, this may seem to be obviously true for the same reasons that $\bp \cdot \p = \strongbp \cdot \bp$. The problem is that amplification requires running some machine which accepts a language in $C$ more than once. This requires the structure of $C$ to allow for such repetitions. This brings us to the concept of majority reductions.
\begin{definition}\label{defmajority}
We say that a complexity class $C$ is \emph{closed under majority reductions} if for any partial function $f \in C$, the partial function $\majority(f)$ defined by
$$\majority(f)(\langle x_1, \ldots, x_k\rangle) = 
\begin{cases}1 & \text{ if the number of $i$ such that }f(x_i) = 1\text{ is at least $\frac{k}{2}$} \\
0 & \text{ if the number of $i$ such that }f(x_i) = 0\text{ is at least $\frac{k}{2}$}\\
\x & \text{ otherwise}
\end{cases}$$
is in $C$.
\end{definition}
Now, we see that majority reductions are exactly what we want for amplification:
\begin{lemma}\label{majorityimpliesamplify}
If $C$ is a complexity class which is closed under majority reductions and polynomial time reductions, then\linebreak \emph{$\bp \cdot C = \strongbp \cdot C$.}
\end{lemma}
\begin{proof}
This is now the same application of the Chernoff bound which shows that $\bp \cdot \p = \strongbp \cdot \p$. Suppose $f$ is a partial function in $\bp \cdot C$. Then there is a partial function $g$ in $C$ such that $\pr_y [g(\langle x,y\rangle) = 1] > 2/3$ if $x\in L$ and $\pr_y [g(\langle x, y\rangle) = 0] > 2/3$ if $x\not\in L$, where the length of $y$ is equal to $p(|x|)$ for some polynomial $p$. Let $q$ be a polynomial, where $e^{-q(|x|)}$ is the desired error probability. Let $k = 48q(|x|)$.

Here is how we determine the value $f(x)$ with error probability $e^{-q(|x|)}$. Choose $k$ strings $y_1,y_2,...,y_k$ randomly, then return $h(\langle x, y_1y_2\cdots y_k\rangle) = \majority(g)(\langle \langle x, y_1 \rangle, \langle x, y_2 \rangle, \ldots, \langle x, y_k \rangle \rangle)$. Notice that $h$ can be easily reduced to $\majority(g)$ in polynomial time, and is therefore in $C$.
Let $X_i$ be an indicator variable which is $1$ if $g(\langle x,y_i\rangle)$ agrees with $f(x)$. We have that the $X_i$ are independent Bernoulli random variables with $\pr\left[X_i = 1\right] > \frac{2}{3}$.  Then by the Chernoff bound,
\begin{align*}
\pr\left[\sum X_i < k/2\right] &= \pr\left[\sum X_i < (1 - 1/4)(2k/3)\right]\\
&< e^{-(1/4)^2 (2k/3) / 2}\\
&= e^{-q(|x|)} \\
&< 2^{-q(|x|)}
\end{align*}
We can do this for any polynomial $q$, so $f \in \strongbp \cdot C$. Hence, $\bp \cdot C = \strongbp \cdot C$. Note that if $f(x) = \x$ then we have no restrictions on what we return.
\end{proof}

@@

To complete the proof of Lemma \ref{amplifymainlemma} (iv), we use the following generalization of a result in \cite{Toda}, which shows that $\parity \cdot (\p ^{\parity\cdot\p}) = \parity\cdot\p$, a fact which is used to show that $\parity \cdot \p$ is closed under majority reductions.

\begin{theorem}\label{oracleparityc}
%Let $C$ be a class containing \emph{$\p$} which is closed under intersection. Then\linebreak \emph{$\parity \cdot (\p^{\parity \cdot C}) = \parity \cdot C$}.
Let $C$ be a class which is closed under intersection. Then \emph{$\parity \cdot (\p^{\parity \cdot C}) = \parity \cdot C$}.
\end{theorem}
\begin{proof}
The direction $\parity \cdot C \subseteq \parity \cdot (\p^{\parity \cdot C})$ is trivial. Let us consider the interesting direction\linebreak $\parity \cdot (\p^{\parity \cdot C})\subseteq \parity\cdot C$.

Suppose $L \in \parity \cdot (P^{\parity\cdot C})$. Then, there is some polynomial time turing machine $T$, equipped with an oracle to some language $M \in \parity\cdot C$, and some polynomial $p$, such that for any $x$, then among all $y$ with $|y| = p(|x|)$, an odd number of them are such that $T$ accepts $(x,y)$, and $T$ rejects on the rest.
\end{proof}
\begin{remark}\emph{
While this paper applies Theorem \ref{oracle} only in the case where $C$ is $\{\Sigma,\Pi,\bp,\parity\}$-constructible, it can also be used to show, for instance, that $\parity \cdot (\p^{\parity \cdot \pp}) = \parity \cdot \pp$, although it is nontrivial to show that $\pp$ is closed under intersection. For a proof, see \cite{Beigel}.
}\end{remark}

>>>>

@@

\begin{theorem}\label{amplify}
Let $C$ be a \emph{$\{\Sigma,\Pi,\bp,\parity\}$}-constructible complexity class. Then \emph{$\bp \cdot C = \strongbp \cdot C$}.
\end{theorem}
Initially, this may seem to be obviously true for the same reasons that $\bp \cdot \p = \strongbp \cdot \bp$. The problem is that amplification requires running some machine which accepts a language in $C$ more than once. This requires the structure of $C$ to allow for such repetitions. This brings us to the concept of majority reductions.
\begin{definition}\label{defmajority}
We say that a complexity class $C$ is \emph{closed under majority reductions} if for any language $L$, the language $\majority(L) = \{(x_1,...,x_k) : k \in \mathbb{N}, |\{x_i : x_i \in L\}| \ge k/2\}$ is in $C$.
\end{definition}
Now, we see that majority reductions are exactly what we want for amplification:
\begin{lemma}\label{majorityimpliesamplify}
If $C$ is a complexity class which is closed under majority reductions, then\linebreak \emph{$\bp \cdot C = \strongbp \cdot C$.}
\end{lemma}
\begin{proof}
This is now the same application of the Chernoff bound which shows that $\bp \cdot \p = \strongbp \cdot \p$. Suppose $L$ is a language in $\bp \cdot C$. Then there is a language $M$ in $C$ such that $\pr_y [(x,y) \in M] > 2/3$ if $x\in L$ and $\pr_y [(x,y)\in M] < 1/3$ if $x\not\in L$, where the length of $y$ is polynomial in the length of $x$. Let $q$ be a polynomial, where $e^{-q(|x|)}$ is the desired error probability. Let $k = 48q(|x|)$.

Here is how we determine if a string $x$ is in $L$ with error probability $e^{-q(|x|)}$. Choose $k$ strings $y_1,y_2,...,y_k$ randomly. Then accept if a majority of $(x,y_1),(x,y_2),...,(x,y_k)$ are in $M$, which is possible to do since $\majority(M) \in C$. Let $X_i$ be an indicator variable which is $1$ if $(x,y_i)$ is in $M$. Then by the Chernoff bound,
\begin{align*}
\pr\left[\sum X_i < k/2\right] &= \pr\left[\sum X_i < (1 - 1/4)(2k/3)\right]\\
&< e^{-(1/4)^2 (2k/3) / 2}\\
&= e^{-q(|x|)} \\
&< 2^{-q(|x|)}
\end{align*}
We can do this for any polynomial $q$, so $L \in \strongbp \cdot C$. Hence, $\bp \cdot C = \strongbp \cdot C$.
\end{proof}

@@

To complete the proof of Lemma \ref{amplifymainlemma} (iv), we use the following generalization of a result in \cite{Toda}, which shows that $\parity \cdot (\p ^{\parity\cdot\p}) = \parity\cdot\p$, a fact which is used to show that $\parity \cdot \p$ is closed under majority reductions. Proving this fact is the main reason that we need the fact that the $\{\Sigma, \Pi, \bp, \parity\}$-constructible classes are closed under intersection.

\begin{theorem}\label{oracleparityc}
%Let $C$ be a class containing \emph{$\p$} which is closed under intersection. Then\linebreak \emph{$\parity \cdot (\p^{\parity \cdot C}) = \parity \cdot C$}.
Let $C$ be a class which is closed under intersection and polynomial time reductions. Then \emph{$ \p^{\parity \cdot C} = \parity \cdot C$}.
\end{theorem}
\begin{proof}
The direction $\parity \cdot C \subseteq \p^{\parity \cdot C}$ is trivial. Let us consider the interesting direction\linebreak $\p^{\parity \cdot C}\subseteq \parity\cdot C$.

Suppose $L \in \parity \cdot (P^{\parity\cdot C})$. Then, there is some polynomial time turing machine $T$, equipped with an oracle to some language $M \in \parity\cdot C$, which accepts $x$ if and only if $x\in L$.

To show that $L$ lies in $\parity\cdot C$, the basic idea is to ``guess" the answers to the oracle calls, and then check them. We make it so that we get an odd number of solutions whenever we guess correctly, and an even number of solutions otherwise.

%We use instead a Turing machine $T'$ which makes oracle calls to the language $M'$, defined as follows. For an input $x$, with $|x| \ge 2$, $M'$ accepts $x$ if either
%\begin{itemize}
%\item $x$ consists only of $0$'s.
%\item $x$ begins with a $1$, and $M$ would accept the last $|x| - 2$ bits.
%\end{itemize}
There exists a polynomial $p$ such that $T$ always makes at most $p(|x|)$ oracle calls. There is a polynomial $q(|x|)$ which bounds the input length of any query made to an oracle.

Each oracle call queries some input to check if it is $M$. There is a language $P \in C$ such that $M$ accepts $x$ if and only if $P$ accepts $(x,y)$ for an odd number of $y$, where $|y| = r(|x|)$ for some polynomial $r$.

To show that $L$ is in $\parity \cdot C$, we guess the following bits for a given input $x$:
\begin{itemize}
\item Bits $a_1,a_2,...,a_{p(|x|)}$. Intuitively, these will be the answers to the oracle calls made by $T$.
\item Bits $b_1,b_2,...,b_{p(|x|)}$. The purpose of these will become clear later.
\item Strings $y_1,y_2,...,y_{p(|x|)}$, each of length $r(q(|x|))$. These are the strings that an oracle call guesses.
\end{itemize}
After guessing, we do the following polynomial time reduction to $C$. For an input $x$, simulate $T$ using the bits $a_1,a_2,...,a_{p(|x|)}$ as the answers to the oracle calls (taking $1$ to mean that the oracle call accepted, and $0$ to mean that the oracle call rejected). If $T$ would reject, then reject.

Also find the strings $x_1,x_2,...,x_{p(|x|)}$ which are the inputs that $T$ would query to the oracles. These may depend on the $a_i$. We now want to verify that the oracle calls are correct. Say that $(a_i, b_i, y_i)$ is \emph{good} if either
\begin{itemize}
\item $b_i = 0$, and both $a_i = 0$ and $y_i = 0^{r(q(|x|))}$, or
\item $b_i = 1$ and $P$ accepts $(x_i, y_i)$
\end{itemize}
Reject if any $(a_i, b_i, y_i)$ is not good. To determine if they are all good, we will have to check some pairs $(x_i, y_i)$ to see if $P$ accepts them. We can do this with a call to $\intersect(P)$, which is in $C$ since $C$ is closed under intersection.

(Technical note: it is possible that some bits out of the $a_i$, $b_i$, and $y_i$ are not used. It is possible that not all $p(x)$ oracle calls will be made, or that not all $r(q(|x|))$ bits of a $y_i$ will be used, if $|x_i| < q(|x|)$. Thus, we also say that $(a_i, b_i, y_i)$ is not good if any unused bits are nonzero.)

Why does this give us what we want? We want to show that this will accept for an odd number of choices of the $a_i$, $b_i$, and $y_i$ if and only if $x \in L$. Let $\# (a_1,...,a_{p(|x|)})$ be the number of ways to choose the $b_i$, and $y_i$ such that all triples $(a_i,b_i,y_i)$ are good. Then the total number of accepting paths is
\begin{center}$\displaystyle \sum_{a_1,...,a_{p(|x|)}\text{ such that $T$ accepts}}  \#(a_1,...,a_{p(|x|)})$\end{center}
We claim that $ \#(a_1,...,a_{p(|x|)})$ is odd if and only if $a_1,a_2,...,a_{p(|x|)}$ are the correct answers to the oracle calls. This follows from the way we chose to decide if $(a_i, b_i, y_i)$ is good. If $a_1,a_2,...,a_{p(|x|)}$ are the correct answers to the oracle calls, then for each $i$, there are an odd number of $(b_i,y_i)$ such that $(a_i, b_i, y_i)$ is good. If $a_i = 1$, all good triples will have $b_i = 1$, so the parity is determined by the number of $y_i$ for which $P$ accepts $(x_i, y_i)$. But if $a_i = 1$, then this should be odd. If $a_i = 0$, then the number of $y_i$ for which $P$ accepts $(x_i, y_i)$ is even. However, we get one more good pair, namely $(0,0,0)$.

On the other hand, if any $a_i$ is incorrect, then the first such incorrect answer will have an even number of good triples.

Hence, the sum will be odd if and only if $T$ accepts when the oracle calls are correct. Therefore, $L$ accepts $x$ if and only if this accepts for an odd number of choices of all the $a_i$, $b_i$, and $y_i$, and $L$ rejects $x$ otherwise. Thus, $L \in \parity \cdot C$. This is true for any $L \in \p^{\parity \cdot C}$, so $\p^{\parity \cdot C} \subseteq \parity \cdot C$.

\end{proof}
\begin{remark}\emph{
While this paper applies Theorem \ref{oracle} only in the case where $C$ is $\{\Sigma,\Pi,\bp,\parity\}$-constructible, it can also be used to show, for instance, that $\parity \cdot (\p^{\parity \cdot \pp}) = \parity \cdot \pp$, although it is nontrivial to show that $\pp$ is closed under intersection. For a proof, see \cite{Beigel}.
}\end{remark}
